---
title: "Aggregate ERA5 Across Healthsheds"
subtitle: "Summarize hourly climate fields into daily boundary-level outputs"
format: html
execute:
  eval: false
---


This page converts raw ERA5 data into daily healthshed-level summaries,
including optional population weighting, timezone adjustment, unit conversion,
and derived variables such as relative humidity and wind speed.

::: {.manual-prereq title="Confirm Aggregation Inputs"}

Before beginning aggregation, confirm that the required boundary geometry,
downloaded ERA5 rasters, elevation data, and optional population rasters are
already available. These are files that may have been staged in a previous workflow step.

```{r}
#| eval: false
shp_file <- "path/to/healthsheds.shp"
data_dir <- "path/to/era5_raw"
elevation_file <- "path/to/elevation_healthshed.csv"
population_dir <- "path/to/worldpop"
output_dir <- "path/to/era5_healthshed_outputs"

list(
  shp_file = shp_file,
  data_dir = data_dir,
  elevation_file = elevation_file,
  population_dir = population_dir,
  output_dir = output_dir
)
```

:::

::: {.manual-procedure title="Declare The Aggregation Configuration"}

The notebook begins by defining whether the run targets ERA5 or ERA5-Land,
which years to process, the local time shift, the healthshed identifier, and
whether values should be population-weighted or area-averaged.

```{r}
#| eval: false
# Pseudocode adapted from the original notebook
is_era5land <- TRUE
id_name <- "fs_uid"
data_years <- 2010:2025
local_time_to_utc <- +3
pop_weighted <- FALSE

source_name <- ifelse(is_era5land, "era5land", "era5")
```

:::

::: {.manual-procedure title="Extract And Clean Hourly Values"}

The core of the notebook does three things for each variable and year:

1. read the raster and align it to the healthshed geometry;
2. extract hourly values per healthshed, optionally weighting by population;
3. clean the hourly series by shifting accumulated variables, adjusting for
   local time, converting units, and applying elevation-based lapse-rate
   corrections where needed.

```{r}
#| eval: false
# Pseudocode adapted from the original notebook
# shp <- read_sf(shp_file)
# r <- rast(file.path(data_dir, "<source>_<variable>_<year>.grib"))
# hourly <- get_hourly_var_per_healthshed(r, year, shp, id_name, pop_weighted = pop_weighted)
# cleaned <- clean_raw_hourly_data(
#   hourly,
#   agg_previous = previous_hours,
#   year = year,
#   variable = variable,
#   is_era5land = is_era5land,
#   local_time_to_utc = local_time_to_utc,
#   elevation_file = elevation_file
# )
```

:::

::: {.manual-procedure title="Write Daily Outputs For Standard Variables"}

For the standard ERA5 variables, the notebook loops over all years, aggregates
cleaned hourly values to daily mean, min, and max, and saves one output file
per variable and year.

```{r}
#| eval: false
# Pseudocode adapted from the original notebook
base_variables <- c(
  "2m_temperature",
  "2m_dewpoint_temperature",
  "surface_pressure",
  "soil_temperature_level_1",
  "volumetric_soil_water_layer_1",
  "volumetric_soil_water_layer_2",
  "total_precipitation",
  "surface_solar_radiation_downwards"
)

# for (variable in base_variables) {
#   for (year in data_years) {
#     daily <- get_daily_avg(cleaned$data_year_clean, id_name = id_name)
#     write.csv(daily, file.path(output_dir, "<source>_<variable>_<year>.csv"))
#   }
# }
```

:::

::: {.manual-procedure title="Derive Relative Humidity And Wind Metrics"}

The notebook also computes additional products that are not downloaded
directly. Relative humidity is derived from temperature and dewpoint; wind
speed and direction are derived from the `u` and `v` wind components.

```{r}
#| eval: false
# Pseudocode adapted from the original notebook
# relative_humidity <- derive_from_temperature_and_dewpoint(hourly_temp, hourly_dewpoint)
# relative_humidity_daily <- get_daily_avg(relative_humidity, round_digits = 1)

# wind_speed <- sqrt(u^2 + v^2)
# wind_speed_daily <- get_daily_avg(wind_speed, round_digits = 2)

# wind_direction <- (180 + atan2(u, v) * 180 / pi) %% 360
# wind_direction_daily <- get_daily_avg(wind_direction, round_digits = 2)
```

:::

::: {.manual-check title="Check The Aggregation Contract"}

A final check section can assert the success of the pipeline.

```{r}
#| eval: false
# This page is complete when the operator can explain:
# - which source files feed the aggregation,
# - when population weighting is used instead of area averaging,
# - why hourly data must be cleaned before daily summaries are computed,
# - and which outputs are direct daily summaries versus derived products
#   such as relative humidity and wind metrics.
```

:::
