---
title: "Orchestrate The ERA5 Pipeline"
subtitle: "Declare the request grid and split it into manageable download jobs"
format: html
execute:
  eval: false
---

::: {.callout-caution}

The content on this page is adapted from a project notebook that mixed pipeline
implementation with explanation. It has been rewritten here as a manual page so
that the workflow intent is easier to inspect step by step.

:::

This page converts the pipeline notebook into an operator-facing setup step.
Its job is to define the overall ERA5 request space and turn it into a set of
grouped API calls that later pages can execute.

::: {.manual-prereq title="Confirm Pipeline Inputs"}

Before building the request grid, confirm that the project already knows where
its staged inputs, stagecoach manifest, and downstream download notebook live.

```{r}
#| eval: false
library(here)

stagecoach_manifest <- here("stagecoach_manifest.yml")
download_era5_nb <- here("notebooks", "download_era5", "download_era5.qmd")
input_dir <- here("data", "inputs")

# In a real manual, these would be concrete prerequisite checks.
stagecoach_manifest
download_era5_nb
input_dir
```

:::

::: {.manual-procedure title="Declare The CDS Request Grid"}

The original notebook defines a large grid of request parameters covering
country, dataset, variable, year, month, day, and hour. It then filters out
invalid dates and avoids redundant hourly requests for time-invariant
variables.

```{r}
#| eval: false
library(tibble)
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)

# Pseudocode adapted from the original pipeline notebook
cds_api_config <- expand_grid(
  years = 2010:2025,
  months = str_pad(1:12, 2, pad = "0"),
  days = 1:31,
  hours = paste0(str_pad(0:23, 2, pad = "0"), ":00"),
  countries = c("MDG", "NPL"),
  datasets = c("reanalysis-era5-land", "reanalysis-era5-single-levels"),
  variables = c(
    "2m_temperature",
    "2m_dewpoint_temperature",
    "volumetric_soil_water_layer_1",
    "volumetric_soil_water_layer_2",
    "surface_solar_radiation_downwards",
    "total_precipitation",
    "10m_u_component_of_wind",
    "10m_v_component_of_wind",
    "surface_pressure",
    "soil_temperature_level_1",
    "geopotential"
  ),
  data_format = "grib",
  download_format = "unarchived"
) |>
  filter(
    !(months %in% c(4, 6, 9, 11) & days == 31),
    !(months == 2 & days > 29),
    !(months == 2 & days == 29 & !leap_year(years)),
    !(variables == "geopotential" & hours != "00:00")
  )
```

:::

::: {.manual-procedure title="Group The Grid Into API Calls"}

The notebook does not send one API call for every row. Instead, it groups the
grid by country, dataset, variable, year, and month, then converts each group
into a single request dictionary that a downstream job can execute.

```{r}
#| eval: false
library(purrr)

# Pseudocode adapted from the original pipeline notebook
request_groups <- cds_api_config |>
  group_by(countries, datasets, variables, years, months) |>
  group_split()

request_dicts <- map(request_groups, \(x) {
  list(
    country = unique(x$countries),
    dataset = unique(x$datasets),
    variable = unique(x$variables),
    year = unique(x$years),
    month = unique(x$months),
    day = unique(x$days),
    time = unique(x$hours),
    data_format = unique(x$data_format),
    download_format = unique(x$download_format)
  )
})
```

:::

::: {.manual-check title="Check The Pipeline Contract"}

```{r}
#| eval: false
# This page is complete when the operator can explain:
# - which inputs are required before downloads begin,
# - which dimensions define the CDS request grid,
# - why invalid dates are filtered out,
# - and why requests are grouped by country, dataset, variable, year, and month.
```

:::
