---
title: "Orchestrate The ERA5 Pipeline"
subtitle: "Declare the request grid and split it into manageable download jobs"
format: html
execute:
  eval: false
---

::: {.callout-caution}

The content on this page contains pseudocode for demonstration purposes.

:::

This page orchestrates the data aggregation pipeline.
Its job is to define the overall ERA5 request space and turn it into a set of
grouped API calls that later pages can execute.

::: {.manual-prereq title="Confirm Pipeline Inputs"}

Before building the request grid, confirm that the project already knows where
its staged inputs, stagecoach manifest, and downstream download notebook live.

```{r}
#| eval: false
library(here)

stagecoach_manifest <- here("stagecoach_manifest.yml")
input_dir <- here("data", "inputs")

# In a real manual, these would be concrete prerequisite checks.
stagecoach_manifest
input_dir
```

:::

::: {.manual-procedure title="Declare The CDS Request Grid"}

Our job defines a large grid of request parameters covering
country, dataset, variable, year, month, day, and hour. It then filters out
invalid dates and avoids redundant hourly requests for time-invariant
variables. End users of a manual could use this space to customize,
tinker with, and experiment with the request grid before it is split into
download jobs, and are encouraged to exercise whatever methods
they prefer to get there, including data transformation libraries,
data visualization tools, and other data analysis frameworks.

```{r}
#| eval: false
library(tibble)
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)

cds_api_config <- expand_grid(
  years = 2010:2025,
  months = str_pad(1:12, 2, pad = "0"),
  days = 1:31,
  hours = paste0(str_pad(0:23, 2, pad = "0"), ":00"),
  countries = c("MDG", "NPL"),
  datasets = c("reanalysis-era5-land", "reanalysis-era5-single-levels"),
  variables = c(
    "2m_temperature",
    "2m_dewpoint_temperature",
    "volumetric_soil_water_layer_1",
    "volumetric_soil_water_layer_2",
    "surface_solar_radiation_downwards",
    "total_precipitation",
    "10m_u_component_of_wind",
    "10m_v_component_of_wind",
    "surface_pressure",
    "soil_temperature_level_1",
    "geopotential"
  ),
  data_format = "grib",
  download_format = "unarchived"
) |>
  filter(
    !(months %in% c(4, 6, 9, 11) & days == 31),
    !(months == 2 & days > 29),
    !(months == 2 & days == 29 & !leap_year(years)),
    !(variables == "geopotential" & hours != "00:00")
  )
```

:::

::: {.manual-procedure title="Group The Grid Into API Calls"}

The notebook does not send one API call for every row. Instead, it groups the
grid by country, dataset, variable, year, and month, then converts each group
into a single request dictionary that a downstream job can execute.

```{r}
#| eval: false
library(purrr)

request_groups <- cds_api_config |>
  group_by(countries, datasets, variables, years, months) |>
  group_split()

request_dicts <- map(request_groups, \(x) {
  list(
    country = unique(x$countries),
    dataset = unique(x$datasets),
    variable = unique(x$variables),
    year = unique(x$years),
    month = unique(x$months),
    day = unique(x$days),
    time = unique(x$hours),
    data_format = unique(x$data_format),
    download_format = unique(x$download_format)
  )
})
```

:::

::: {.manual-check title="Check The Pipeline Contract"}
In some cases, authors may choose not to include a file artifact
in a check block, but instead an end-of-page summary that 
describes what the operator should be able to explain or demonstrate.

```{r}
#| eval: false
# This page is complete when the operator can explain:
# - which inputs are required before downloads begin,
# - which dimensions define the CDS request grid,
# - why invalid dates are filtered out,
# - and why requests are grouped by country, dataset, variable, year, and month.
stopifnot({
  "I can explain the required inputs, the request grid dimensions, the filtering of invalid dates, and the grouping of requests."
})
```

:::
