---
title: Quarto Manuals
author: Edenian
subtitle: Executable software operations manuals built from independent Quarto pages
header_font: "Baskerville"
header_logo: "dummy-thchan.png"
header_block_color: "A51C30"
heading_color: "1E1E1E"
body_color: "F4F4F4"
body_font: "Arial"
format:
  poster-typst: 
    keep-typ: true
execute:
  eval: true
---

PLACEHOLDER CONTENT. NOT YET IMPLEMENTED

```{r}
#| echo: false
library(ggplot2)
var <- "JN.1"
today_date <- Sys.Date()
```

```{=typst}
#import "@preview/showybox:2.0.4": showybox 
#let main = rgb("#A51C30")

#showybox(
  title-style: (
    boxed-style: (
      anchor: (
        x: center,
        y: horizon
      ),
      radius: (top-left: 10pt, bottom-right: 10pt, rest: 0pt),
    )
  ),
  frame: (
    title-color: main,
    body-color: main.lighten(80%),
    footer-color: main.lighten(60%),
    border-color: main.darken(20%),
    radius: (top-left: 10pt, bottom-right: 10pt, rest: 0pt)

  ),
  title: "tl;dr",
  footer: text(size: 30pt, weight: 600, "Check it out: you can also add Typst code to make math functions or diagrams.")
)[
  This is a Typst template that can be used with Quarto to execute code and automatically render outputs like text, plots, and tables.
  Forget about copying and pasting screenshots 

  $ frac(partial^n f, partial x_i ... partial x_j)(a_1, a_2, ..., a_n) = frac(partial^n f, partial x_j ... partial x_i)(a_1, a_2, ..., a_n) $
]
```

:::{.block fill="white" inset="30pt" radius="6pt"}

# **INSTALLATION**

To install this poster template, go to your project root and run:

```bash
quarto install extension https://codeberg.org/edenian-prince/quarto-typst-poster/archive/main.zip
```

You can also add the template/demo quarto document by running the code below,
but note, it adds all the files in the repo currently..

```bash
quarto use template https://codeberg.org/edenian-prince/quarto-typst-poster/archive/main.zip
```

:::

::::{.block fill="white" inset="30pt" radius="6pt"}


# **USAGE**

After installing, change the format in your Quarto document's yaml front matter to `format: poster-typst`

```markdown
---
format:
  poster-typst: default
---
```

You can also add other options in the front matter like a title, subtitle, and color options. See the codeberg repo for a complete list of options.


# **ADDING SECTIONS**

This format is a Typst canvas, but Quarto is where you write the sections that contain text, code, and outputs. These sections are what you are 
reading now. To create these sections in white blocks (or whatever color you want...), wrap your sections in fenced divs like this:

````markdown
:::{.block fill="white" inset="30pt" radius="6pt"}

Headings, text, and code go here

:::
````


::::



:::{.block fill="white" inset="30pt" radius="6pt"}

# **ADDING PLOTS**

Run put your code in the Quarto document like you normally would and it will render here.

- Center the plot with `#| fig-align: center` 
- Adjust the width with `#| fig-width: 12`
- Adjust the height with `#| fig-height: 9`
- Adjust plot attributes in your code

```{r}
#| message: false 
#| warning: false
#| echo: false 
#| fig-width: 12
#| fig-height: 9
#| fig-align: center 
mpg2 <- data.frame(
  name = c("Sedan", "SUV", "Truck", "Hybrid", "Electric"),
  value = c(25, 18, 15, 45, 120)
)
data <- data.frame(x=rnorm(100))

# Custom a few word of the title only:
my_title <- expression(paste("A ", italic(bold("blue")), " histogram"))
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.8) +
  ggtitle(my_title) +
  theme_minimal() +
  theme(
        plot.title = element_text(size=38),
        axis.title.x = element_text(size=30),
        axis.text.x = element_text(size=30),
        axis.title.y = element_text(size=30),
        axis.text.y = element_text(size=30)
  )


```
Add a footer or something underneath

:::

:::{.block fill="white" inset="30pt" radius="6pt"}

# **ADDING TABLES**

Likewise, you can add tables by doing the same as above.


```{r}
#| message: false 
#| warning: false
#| echo: false 
#| fig-width: 12
#| fig-height: 9
#| fig-align: center 
library(gtsummary)
# Setting JAMA theme for gtsummary
theme_gtsummary_journal("jama")
#> Setting theme "JAMA"
# Themes can be combined by including more than one
theme_gtsummary_compact()
#> Setting theme "Compact"

gt_table <- trial |>
  select(age, grade, trt) |>
  tbl_summary(by = trt) |>
  as_gt() |>
  gt::tab_options(
    table.font.size = gt::px(30)
  )

library(tinytable)
library(gt)
x <- gtcars |>
  select(1:3) |> 
  head() 

x |>
  tt(width = c(.25, .55, .25)) |>
  style_tt(align = c("lcc")) |> 
  style_tt(
    j = 3, 
    i = which(x$year == 2015),
    color = "red"
)
```

`tinytable` seems to look the best in this format fyi.

:::



:::{.block fill="white" inset="30pt" radius="6pt"}

# **TYPST CODE**

You can add Typst snippets in the Quarto doc like this with showybox.
This is how I made the `tl;dr` and `Source Code` blocks.

````markdown
```{=typst}
#import "@preview/showybox:2.0.4": showybox 

#showybox(...)

```
````

:::

:::{.block fill="white" inset="30pt" radius="6pt"}

# **RENDER THE POSTER**

````bash 
quarto render poster.qmd
````

:::



```{=typst}

#showybox(
  title-style: (
    boxed-style: (
      anchor: (
        x: center,
        y: horizon
      ),
      radius: (top-left: 10pt, bottom-right: 10pt, rest: 0pt),
    )
  ),
  frame: (
    title-color: main,
    body-color: main.lighten(80%),
    footer-color: main.lighten(60%),
    border-color: main.darken(20%),
    radius: (top-left: 10pt, bottom-right: 10pt, rest: 0pt)
  ),
  title: "Source Code",
)[
```
<!-- im adding the qr code between the typst block here, fyi -->

```{r}
#| echo: false
#| fig-align: center 
#| fig-height: 5
#| fig-wigth: 6
code <- qrcode::qr_code("https://codeberg.org/edenian-prince/quarto-typst-poster ")
plot(code)
```

```{=typst}
]
```



